---
title: "bootstrap"
output: html_document
---

## Looking at bootstrap distribution

```{r}
set.seed(333); x <- rnorm(30)
bootMean <- rep(NA,1000); sampledMean <- rep(NA,1000)
for(i in 1:1000){bootMean[i] <- mean(sample(x,replace=TRUE))}
for(i in 1:1000){sampledMean[i] <- mean(rnorm(30))}
plot(density(bootMean)); lines(density(sampledMean),col="red")
```

## Example with boot package

```{r}
library(boot)
set.seed(333); x <- rnorm(30); sampledMean <- rep(NA,1000)
for(i in 1:1000){sampledMean[i] <- mean(rnorm(30))}
meanFunc <- function(x,i){mean(x[i])}
bootMean <- boot(x,meanFunc,1000)
bootMean

```

Now we plot it

```{r}
plot(density(bootMean$t)); lines(density(sampledMean),col="red")
```





## Bootstrap regression

```{r}
library(boot); data(nuclear)
nuke.lm <- lm(log(cost) ~ date,data=nuclear)
plot(nuclear$date,log(nuclear$cost),pch=19)
abline(nuke.lm,col="red",lwd=3)
```

Now bootstrap it

```{r}
par(mfrow=c(1,3))
for(i in 1:3){
  nuclear0 <- nuclear[sample(1:dim(nuclear)[1],replace=TRUE),]
  nuke.lm0 <- lm(log(cost) ~ date,data=nuclear0)
  plot(nuclear0$date,log(nuclear0$cost),pch=19)
  abline(nuke.lm0,col="red",lwd=3)
}
```

Bootstrap distribution

```{r}
bs <- function(data, indices,formula) {
  d <- data[indices,];fit <- lm(formula, data=d);return(coef(fit)) 
} 
results <- boot(data=nuclear, statistic=bs, R=1000, formula=log(cost) ~ date)
plot(density(results$t[,2]),col="red",lwd=3)
lines(rep(nuke.lm$coeff[2],10),seq(0,8,length=10),col="blue",lwd=3)
```


## modelr example

```{r}
library(broom)
library(purrr)
library(modelr)
library(tidyr)

pboot <- nuclear %>%
  bootstrap(1000) %>%
  mutate(mod = map(strap, ~ lm(log(cost) ~ date,data=.)))  %>%
  unnest(map(mod, tidy)) %>%
  filter(term=="date")

plot(density(pboot$estimate),col="red",lwd=3)
lines(rep(nuke.lm$coeff[2],10),seq(0,8,length=10),col="blue",lwd=3)
```


## A few ways to bootstrap a regression model

### Bootstrapping samples - to get a variance

```{r}
f1 = formula(log(cost) ~ date)
lm1 = lm(f1,data=nuclear)
stat = lm1$coeff[2]

n = dim(nuclear)[1]
B = 500
stat0 = rep(0,B)

set.seed(1262016)

for(i in 1:B){
  nuclear0 = nuclear[sample(1:n,replace=T),]
  lm0 = lm(f1,data=nuclear0)
  stat0[i] = lm0$coeff[2]
}
hist(stat0)
abline(v=stat,col="blue")
```


### Bootstrapping residuals - to get a variance

```{r}
f1 = formula(log(cost) ~ date)
lm1 = lm(f1,data=nuclear)
lm_null = lm(log(cost) ~ 1,data=nuclear)
stat = lm1$coeff[2]
resid = lm1$residuals
null_fit = lm_null$fitted.values

n = dim(nuclear)[1]
B = 500
stat0 = rep(0,B)

set.seed(1262016)

for(i in 1:B){
  resid0 = resid[sample(1:n,replace=T)]
  nuclear$y0 = null_fit + resid0
  lm0 = lm(y0 ~ date,data=nuclear)
  stat0[i] = lm0$coeff[2]
}
hist(stat0,xlim=c(-0.2,0.3))
abline(v=stat,col="blue")
```



### Bootstrapping residuals parametrically -   to get a variance

```{r}
f1 = formula(log(cost) ~ date)
lm1 = lm(f1,data=nuclear)
stat = lm1$coeff[2]
resid_sd = sd(lm1$residuals)
fit = lm1$fitted.values

n = dim(nuclear)[1]
B = 500
stat0 = rep(0,B)

set.seed(1262016)

for(i in 1:B){
  nuclear$y0 = rnorm(n,mean=fit,sd=resid_sd)
  lm0 = lm(y0 ~ date,data=nuclear)
  stat0[i] = lm0$coeff[2]
}
hist(stat0)
abline(v=stat,col="blue")
```


### Bootstrapping residuals non-parametrically - to test a hypothesis

```{r}
f1 = formula(log(cost) ~ date)
lm1 = lm(f1,data=nuclear)
stat = lm1$coeff[2]
resid_sd = sd(lm1$residuals)
fit = lm1$fitted.values

n = dim(nuclear)[1]
B = 500
stat0 = rep(0,B)

set.seed(1262016)

for(i in 1:B){
  nuclear$y0 = rnorm(n,mean=fit,sd=resid_sd)
  lm0 = lm(y0 ~ date,data=nuclear)
  stat0[i] = lm0$coeff[2]
}
hist(stat0)
abline(v=stat,col="blue")
```


## Stamps example

```{r}
library(bootstrap)
data(stamp)
plot(density(stamp[,1]))
```

## Test if there are more than 2 modes?

Can you create a bootstrap hypothesis test that calculates a p-value for their being more than 2 modes. 


