---
title: "Spam example"
output: html_document
---

# Example of how to build a predictor in R


## Question: Can we predict spam emails from the words in the emails? 

Dear Jeff,

Can you send me your address so I can send you the invitation?

Thanks,

Ben

### Load the data

```{r}
library(kernlab)
data(spam)
head(spam)
```


### Training and test sets

```{r}
set.seed(111416)
n = dim(spam)[1]
index = sample(1:n,size=n*0.75,replace=F)
training=spam[index,]
test = spam[(1:n)[-index],]

```


## Features - potential predictor frequency of "your"

The frequency of "your" in this email is: 

$$ 2/17 = 0.118$$

### Look at potential predictor

```{r}
plot(density(training$your[training$type=="nonspam"]),
     col="blue",main="",xlab="Frequency of 'your'")
lines(density(training$your[training$type=="spam"]),col="red")

```

## Algorithm - constant cutoff

```{r}
prediction <- ifelse(training$your > 0.5,"spam","nonspam")

```


## Evaluation 

### In sample

```{r}
table(prediction,training$type)/length(training$type)
```

```{r}
mean(prediction==training$type)
```

### Could adjust algorithm here

```{r}
prediction2 <- ifelse(training$your > 0.6,"spam","nonspam")
```


### Out of sample

Only our best predictor here

```{r}
testpredict = ifelse(test$your > 0.6, "spam","nonspam")
table(testpredict,test$type)/length(test$type)
mean(testpredict==test$type)
```


# Example of overfitting

## A really small data set

```{r}
library(kernlab); data(spam); set.seed(333)
smallSpam <- spam[sample(dim(spam)[1],size=10),]
spamLabel <- (smallSpam$type=="spam")*1 + 1
plot(smallSpam$capitalAve,col=spamLabel)
```

## Predictor 1

* capitalAve $>$ 2.7 = "spam"
* capitalAve $<$ 2.40 = "nonspam"
* capitalAve between 2.40 and 2.45 = "spam"
* capitalAve between 2.45 and 2.7 = "nonspam"


## Application

```{r}
rule1 <- function(x){
  prediction <- rep(NA,length(x))
  prediction[x > 2.7] <- "spam"
  prediction[x < 2.40] <- "nonspam"
  prediction[(x >= 2.40 & x <= 2.45)] <- "spam"
  prediction[(x > 2.45 & x <= 2.70)] <- "nonspam"
  return(prediction)
}
table(rule1(smallSpam$capitalAve),smallSpam$type)
```


## Predictor 2


* capitalAve $>$ 2.40 = "spam"
* capitalAve $\leq$ 2.40 = "nonspam"



## Application

```{r}

rule2 <- function(x){
  prediction <- rep(NA,length(x))
  prediction[x > 2.8] <- "spam"
  prediction[x <= 2.8] <- "nonspam"
  return(prediction)
}
table(rule2(smallSpam$capitalAve),smallSpam$type)

```



## Apply to spam data

Look at predictions

```{r}

table(rule1(spam$capitalAve),spam$type)
table(rule2(spam$capitalAve),spam$type)
```


Accuracy

```{r}
mean(rule1(spam$capitalAve)==spam$type)
mean(rule2(spam$capitalAve)==spam$type)
```



