---
title: "Caret spam example"
output: html_document
---


## SPAM Example: Data splitting

```{r }
library(caret); library(kernlab); data(spam)
inTrain <- createDataPartition(y=spam$type,
                              p=0.75, list=FALSE)
training <- spam[inTrain,]
testing <- spam[-inTrain,]
dim(training)
```


 

## SPAM Example: Fit a model

```{r }
set.seed(32343)
modelFit <- train(type ~.,data=training, method="glm")
modelFit
```


 

## SPAM Example: Final model

```{r }
modelFit$finalModel
```


 

## SPAM Example: Prediction

```{r predictions, dependson="training",cache=TRUE}
predictions <- predict(modelFit,newdata=testing)
predictions
```

 

## SPAM Example: Confusion Matrix

```{r confusion, dependson="predictions",cache=TRUE}
confusionMatrix(predictions,testing$type)
```



## SPAM Example: K-fold

```{r }
set.seed(32323)
folds <- createFolds(y=spam$type,k=10,
                             list=TRUE,returnTrain=TRUE)
sapply(folds,length)
folds[[1]][1:10]
```



## SPAM Example: Return test

```{r }
set.seed(32323)
folds <- createFolds(y=spam$type,k=10,
                             list=TRUE,returnTrain=FALSE)
sapply(folds,length)
folds[[1]][1:10]
```



## SPAM Example: Resampling

```{r }
set.seed(32323)
folds <- createResample(y=spam$type,times=10,
                             list=TRUE)
sapply(folds,length)
folds[[1]][1:10]
```



## SPAM Example: Time Slices {.smaller}

```{r }
set.seed(32323)
tme <- 1:1000
folds <- createTimeSlices(y=tme,initialWindow=20,
                          horizon=10)
names(folds)
folds$train[[1]]
folds$test[[1]]
```



```{r }
library(caret); library(kernlab); data(spam)
inTrain <- createDataPartition(y=spam$type,
                              p=0.75, list=FALSE)
training <- spam[inTrain,]
testing <- spam[-inTrain,]
hist(training$capitalAve,main="",xlab="ave. capital run length")
```



## Why preprocess?

```{r }
mean(training$capitalAve)
sd(training$capitalAve)
```



## Standardizing

```{r }
trainCapAve <- training$capitalAve
trainCapAveS <- (trainCapAve  - mean(trainCapAve))/sd(trainCapAve) 
mean(trainCapAveS)
sd(trainCapAveS)
```



## Standardizing - test set

```{r }
testCapAve <- testing$capitalAve
testCapAveS <- (testCapAve  - mean(trainCapAve))/sd(trainCapAve) 
mean(testCapAveS)
sd(testCapAveS)
```




## Standardizing - _preProcess_ function

```{r }
preObj <- preProcess(training[,-58],method=c("center","scale"))
trainCapAveS <- predict(preObj,training[,-58])$capitalAve
mean(trainCapAveS)
sd(trainCapAveS)
```




## Standardizing - _preProcess_ function

```{r }
testCapAveS <- predict(preObj,testing[,-58])$capitalAve
mean(testCapAveS)
sd(testCapAveS)
```



## Standardizing - _preProcess_ argument

```{r }
set.seed(32343)
modelFit <- train(type ~.,data=training,
                  preProcess=c("center","scale"),method="glm")
modelFit
```




## Standardizing - Box-Cox transforms

```{r }
preObj <- preProcess(training[,-58],method=c("BoxCox"))
trainCapAveS <- predict(preObj,training[,-58])$capitalAve
par(mfrow=c(1,2)); hist(trainCapAveS); qqnorm(trainCapAveS)
```




## Standardizing - Imputing data

```{r }
set.seed(13343)

# Make some values NA
training$capAve <- training$capitalAve
selectNA <- rbinom(dim(training)[1],size=1,prob=0.05)==1
training$capAve[selectNA] <- NA

# Impute and standardize
preObj <- preProcess(training[,-58],method="knnImpute")
capAve <- predict(preObj,training[,-58])$capAve

# Standardize true values
capAveTruth <- training$capitalAve
capAveTruth <- (capAveTruth-mean(capAveTruth))/sd(capAveTruth)
```




## Standardizing - Imputing data

```{r ,dependson="knn",cache=TRUE,fig.height=3.5,fig.width=7}
quantile(capAve - capAveTruth)
quantile((capAve - capAveTruth)[selectNA])
quantile((capAve - capAveTruth)[!selectNA])
```



## Correlated predictors

```{r }
library(caret); library(kernlab); data(spam)
inTrain <- createDataPartition(y=spam$type,
                              p=0.75, list=FALSE)
training <- spam[inTrain,]
testing <- spam[-inTrain,]

M <- abs(cor(training[,-58]))
diag(M) <- 0
which(M > 0.8,arr.ind=T)
```



## Correlated predictors

```{r }
names(spam)[c(34,32)]
plot(spam[,34],spam[,32])
```


## Principal components in R - prcomp

```{r }
smallSpam <- spam[,c(34,32)]
prComp <- prcomp(smallSpam)
plot(prComp$x[,1],prComp$x[,2])
```



## Principal components in R - prcomp

```{r }
prComp$rotation
```




## PCA on SPAM data

```{r }
typeColor <- ((spam$type=="spam")*1 + 1)
prComp <- prcomp(log10(spam[,-58]+1))
plot(prComp$x[,1],prComp$x[,2],col=typeColor,xlab="PC1",ylab="PC2")
```




## PCA with caret

```{r }
preProc <- preProcess(log10(spam[,-58]+1),method="pca",pcaComp=2)
spamPC <- predict(preProc,log10(spam[,-58]+1))
plot(spamPC[,1],spamPC[,2],col=typeColor)
```




## Preprocessing with PCA

```{r }
preProc <- preProcess(log10(training[,-58]+1),method="pca",pcaComp=2)
trainPC <- predict(preProc,log10(training[,-58]+1))
trainPC <- cbind(trainPC,type=training$type)
modelFit <- train(type ~ PC1 + PC2,method="glm",data=trainPC)
```



## Preprocessing with PCA

```{r}
testPC <- predict(preProc,log10(testing[,-58]+1))
confusionMatrix(testing$type,predict(modelFit,testPC))
```



## Alternative (sets # of PCs)

```{r }
modelFit <- train(training$type ~ .,method="glm",preProcess="pca",data=training)
confusionMatrix(testing$type,predict(modelFit,testing))
```






## Trees

```{r }
data(iris); library(ggplot2)
names(iris)
table(iris$Species)
```


## Training/testing

```{r }
inTrain <- createDataPartition(y=iris$Species,
                              p=0.7, list=FALSE)
training <- iris[inTrain,]
testing <- iris[-inTrain,]
dim(training); dim(testing)
```

## Plot predictors

```{r}
qplot(Petal.Width,Sepal.Width,colour=Species,data=training)
```

## Fit a tree

```{r}
library(caret)
modFit <- train(Species ~ .,method="rpart",data=training)
print(modFit$finalModel)
```

## Plot tree

```{r}
plot(modFit$finalModel, uniform=TRUE, 
      main="Classification Tree")
text(modFit$finalModel, use.n=TRUE, all=TRUE, cex=.8)
```

## Predicting new values

```{r}
predict(modFit,newdata=testing)
```



## Ozone data

```{r}
library(ElemStatLearn); data(ozone,package="ElemStatLearn")
ozone <- ozone[order(ozone$ozone),]
head(ozone)
```



## Bagged loess
```{r}
ll <- matrix(NA,nrow=10,ncol=155)
for(i in 1:10){
  ss <- sample(1:dim(ozone)[1],replace=T)
  ozone0 <- ozone[ss,]; ozone0 <- ozone0[order(ozone0$ozone),]
  loess0 <- loess(temperature ~ ozone,data=ozone0,span=0.2)
  ll[i,] <- predict(loess0,newdata=data.frame(ozone=1:155))
}
```


## Plot it

```{r}
plot(ozone$ozone,ozone$temperature,pch=19,cex=0.5)
for(i in 1:10){lines(1:155,ll[i,],col="grey",lwd=2)}
lines(1:155,apply(ll,2,mean),col="red",lwd=2)
```


## Back to the iris data

```{r}
data(iris); library(ggplot2)
inTrain <- createDataPartition(y=iris$Species,
                              p=0.7, list=FALSE)
training <- iris[inTrain,]
testing <- iris[-inTrain,]
```


## Fit a random forest

```{r}
library(caret)
modFit <- train(Species~ .,data=training,method="rf",prox=TRUE)
modFit
```

## Look at a single tree

```{r}
getTree(modFit$finalModel,k=2)
```

## Plot class "centers"

```{r}
irisP <- classCenter(training[,c(3,4)], training$Species, modFit$finalModel$prox)
irisP <- as.data.frame(irisP); irisP$Species <- rownames(irisP)
p <- qplot(Petal.Width, Petal.Length, col=Species,data=training)
p + geom_point(aes(x=Petal.Width,y=Petal.Length,col=Species),size=5,shape=4,data=irisP)
```

## Predictng new values

```{r}
pred <- predict(modFit,testing); testing$predRight <- pred==testing$Species
table(pred,testing$Species)
```

## Plot predictions

```{r}
qplot(Petal.Width,Petal.Length,colour=predRight,data=testing,main="newdata Predictions")
```

## Variable importance

```{r}
varImpPlot(modFit$finalModel)
```


## Boosting, wage example

```{r}
library(ISLR); data(Wage); library(ggplot2); library(caret);
Wage <- subset(Wage,select=-c(logwage))
inTrain <- createDataPartition(y=Wage$wage,
                              p=0.7, list=FALSE)
training <- Wage[inTrain,]; testing <- Wage[-inTrain,]
```


## Fit the model

```{r}
modFit <- train(wage ~ ., method="gbm",data=training,verbose=FALSE)
print(modFit)
```


## Plot the results

```{r}
qplot(predict(modFit,testing),wage,data=testing)
```


## Visualize what the model is doing

```{r}
plot(modFit$finalModel,1)
plot(modFit$finalModel,2)
```




