---
title: "Spam example"
output: html_document
---

# Example of how to build a predictor in R


## Question: Can we predict spam emails from the words in the emails? 

Dear Jeff,

Can you send me your address so I can send you the invitation?

Thanks,

Ben

### Load the data

```{r}
library(kernlab)
data(spam)
head(spam)
```


### Training and test sets

```{r}
set.seed(111416)
n = dim(spam)[1]
index = sample(1:n,size=n*0.75,replace=F)
training=spam[index,]
test = spam[(1:n)[-index],]

```


## Features - potential predictor frequency of "your"

The frequency of "your" in this email is: 

$$ 2/17 = 0.118$$

### Look at potential predictor

```{r}
plot(density(training$your[training$type=="nonspam"]),
     col="blue",main="",xlab="Frequency of 'your'")
lines(density(training$your[training$type=="spam"]),col="red")

```

## Algorithm - constant cutoff

```{r}
prediction <- ifelse(training$your > 0.5,"spam","nonspam")

```


## Evaluation 

### In sample

```{r}
table(prediction,training$type)/length(training$type)
```

```{r}
mean(prediction==training$type)
```

### Could adjust algorithm here

```{r}
prediction2 <- ifelse(training$your > 0.6,"spam","nonspam")
```


### Out of sample

Only our best predictor here

```{r}
testpredict = ifelse(test$your > 0.6, "spam","nonspam")
table(testpredict,test$type)/length(test$type)
mean(testpredict==test$type)
```
