---
title: "EDA lab"
author: "Jeff Leek"
date: "October 17, 2016"
output: html_document
---


1.  Make a plot of latitude versus longitude for the quakes data in base R. 

```{r}
plot(quakes$lat,quakes$long)
```

2. Make the points filled in circles

```{r}
plot(quakes$lat,quakes$long,pch=19)
```

2. Color the figure by the number of stations reporting (hint use the `cut2` function). 

```{r}
library(Hmisc)
sfactor = cut2(quakes$stations)
table(sfactor,as.numeric(sfactor))
cols=colorRampPalette(c("lightblue","darkblue"))(6)

plot(quakes$lat,quakes$long,col=cols[as.numeric(sfactor)],pch=19)
```

3. Make the axis labels "Latitude" and "Longitude"

```{r}
plot(quakes$lat,quakes$long,
     col=cols[as.numeric(sfactor)],
     pch=19,
     xlab="Latitude",ylab="Longitude")
```

3. Make the axis labels and ticks larger

```{r}
plot(quakes$lat,quakes$long,
     col=cols[as.numeric(sfactor)],
     pch=19,
     xlab="Latitude",ylab="Longitude",cex.lab=1.5,cex.axis=1.5)
```

4. Add the title "Spatial Distribution of Stations"

```{r}
plot(quakes$lat,quakes$long,
     col=cols[as.numeric(sfactor)],
     pch=19,
     xlab="Latitude",ylab="Longitude",
     cex.lab=1.5,cex.axis=1.5,
     main="Spatial Distribution of Stations")
```


5. Add a legend showing the coloring

```{r}
plot(quakes$lat,quakes$long,
     col=cols[as.numeric(sfactor)],
     pch=19,
     xlab="Latitude",ylab="Longitude",
     cex.lab=1.5,cex.axis=1.5,
     main="Spatial Distribution of Stations")
legend(-35,175,legend=unique(sfactor),
       col=cols[1:6],pch=19)
```

6. Add a title to the legend

```{r}
plot(quakes$lat,quakes$long,
     col=cols[as.numeric(sfactor)],
     pch=19,
     xlab="Latitude",ylab="Longitude",
     cex.lab=1.5,cex.axis=1.5,
     main="Spatial Distribution of Stations")
legend(-35,175,legend=unique(sfactor),
       col=cols[1:6],pch=19)
text(-31,177,"Number\n of stations\n reporting")
```


